---
title: "Untitled"
author: "Your Name Here"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

setwd("C:/Users/eschlatter/Dropbox/DispersalSGS/analysis")
#treefile = "../output/ts_8238198510834684392_t160000"
```

Load empirical FST
```{r}
fst_emp <- read.csv('../data/FST_empirical.csv',header=T)
fst_emp_long <- pivot_longer(fst_emp,cols=2:ncol(fst_emp),names_to=c('site2'))
fst_emp_long$sites <- NA
for(i in 1:nrow(fst_emp_long)){
  fst_emp_long$sites[i] <- paste(sort(as.matrix(fst_emp_long[i,1:2])),collapse='.')
}
fst_emp_long <- filter(fst_emp_long,!is.na(value)) %>%
  rename(emp_fst=value)
fst_emp <- fst_emp_long
rm(fst_emp_long)
```

Load simulated FSTs
```{r}
files <- list.files(path='../output/FSTs_240k') %>% strsplit(" ")
fst_sims <- list()
for (i in 1:length(files)) {
   load(paste0('../output/FSTs_240k/',files[[i]])) # load fst matrix
  
   fst_sim <- as.data.frame(fsts) %>%
     rownames_to_column(var='site1')
   fst_sim <- pivot_longer(fst_sim,cols=2:ncol(fst_sim),names_to=c('site2'))
   
   fst_sim$sites <- NA
   for(j in 1:nrow(fst_sim)){
     fst_sim$sites[j] <- paste(sort(as.matrix(fst_sim[j,1:2])),collapse='.')
   }
   fst_sim <- filter(fst_sim,!is.na(value)) %>%
     rename(sim_fst=value)
     #rename_with(~paste0('sim_fst_',i),value)
   
   assign(paste0('fst_sim_',i), fst_sim) # save as "fst_sim_i"
   fst_sims[[i]] <- fst_sim
}
```

Combine simulated and empirical
```{r}
# combine simulation runs into a long table, with a column for the simulation number
all_fst_sim <- data.frame(site1=factor(), site2=factor(), sim_fst=numeric(), sites=factor(), sim=factor())
for(i in 1:length(fst_sims)){
  this_sim <- fst_sims[[i]]
  all_fst_sim <- rbind(all_fst_sim,mutate(this_sim, sim=as.factor(i)))
}

# add empirical data as its own column
all_fst_both <- full_join(all_fst_sim,fst_emp,by='sites') %>%
  select(sites,site1,site2=site2.x,sim,sim_fst,emp_fst) %>%
  filter(!is.na(emp_fst))

# group by site pair (across simulations)
all_fst_group <- group_by(all_fst_both,sites) %>%
  summarize(emp_fst=first(emp_fst),site1=first(site1),site2=first(site2),
            sim_med=median(sim_fst),sim_sd=sd(sim_fst),sim_min=min(sim_fst),sim_max=max(sim_fst))

# add extra columns for site, population, and population pair, to make exploratory plotting easier
site_list <- unique(c(fst_emp$X,fst_emp$site2))
site_list <- c(site_list[length(site_list)],site_list[1:(length(site_list)-1)]) # put them in order
for(i in site_list){
  all_fst_group <- mutate(all_fst_group,newsite=(i==site1 | i==site2)) %>%
    rename_with(~paste0(i),newsite)
}
for(i in c('B','T','G','L')){
  all_fst_group <- mutate(all_fst_group,newsite=grepl(i,sites)) %>%
    rename_with(~paste0(i),newsite)
}
all_fst_group <- mutate(all_fst_group,first_pop=substr(sites,1,1)) %>%
  mutate(second_pop=substr(sub('.*\\.','',sites),1,1)) %>%
  mutate(both_pops=paste0(first_pop,second_pop))

save(all_fst_group,file='../output/FST_all_240k.RData')
```


```{r}
load('../output/FST_all_300k.RData')

ggplot(all_fst_group,aes(x=emp_fst,y=sim_med))+
  geom_point()+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')
```


```{r}
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_point()+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')+
  geom_smooth(method='lm',se=FALSE)+
  ggtitle('300k generations')
```

```{r}
ggplot(filter(all_fst_group,both_pops %in% c('TT','GG','LL')),aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_pointrange(aes(ymin=sim_min,ymax=sim_max,color=both_pops),position=position_jitter(width=0.0001))+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')+
  geom_smooth(method='lm',se=FALSE)
```


Looks like there's something going on with the Barrier-Lighthouse group. Some simulated values are much larger than others, for similar empirical values.
First check: is it just among-sim variation?
```{r}
BL <- filter(all_fst_group,both_pops=="BL")# %>% select(sites,emp_fst,sim_med,sim_sd,sim_min,sim_max)
ggplot(BL,aes(x=emp_fst,y=sim_med))+
  geom_pointrange(aes(ymin=sim_med-sim_sd,ymax=sim_med+sim_sd),
                  position=position_jitter(width=0.0008))+
  geom_point(data=all_fst_group,aes(x=emp_fst,y=sim_med),color='gray')
```
Nope. There are definitely some points that are distinct from the rest. Figure out which ones they are:
It's B1, B2, and B3 with Lighthouse.
```{r}
filter(BL,sim_med>0.175)

ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,color=(L & (B1 | B2 | B3 | B5))))+
  geom_point()
```

```{r}
## highlight a single point
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med))+
  geom_point(size=1)+
  geom_point(data=filter(all_fst_group,B10==TRUE & T6==TRUE),aes(x=emp_fst,y=sim_med),color='red')
```



```{r}
## highlight the points from one site
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,fill=both_pops,color=both_pops),shape=21)+
  geom_point(size=1.1)+
  geom_smooth(method='lm',se=FALSE)+
  geom_point(data=filter(all_fst_group,T6==TRUE),aes(x=emp_fst,y=sim_med,fill=both_pops),color='black',shape=21)+
  ggtitle('T6')
```
```{r}
## highlight the points from one site
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,fill=both_pops,color=both_pops),shape=21)+
  geom_point(size=1.1)+
  geom_smooth(method='lm',se=FALSE)+
  geom_point(data=filter(all_fst_group,(B1==TRUE | B2==TRUE | B3==TRUE)),aes(x=emp_fst,y=sim_med,fill=both_pops),color='black',shape=21)+
  ggtitle('B1,B2,B3')
```


```{r}
hilite_site="B20"
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,fill=both_pops,color=both_pops),shape=21)+
  geom_point(size=1.1)+
  geom_smooth(method='lm',se=FALSE)+
  geom_point(data=filter(all_fst_group,B20==TRUE),aes(x=emp_fst,y=sim_med,fill=both_pops),color='black',shape=21)+
  ggtitle(hilite_site)
```


```{r}
bigs <- filter(all_fst_group,sim_med>0.2) %>% select(sites,emp_fst,sim_med,sim_sd)
ggplot(bigs,aes(x=emp_fst,y=sim_med))+
  geom_point()+
  geom_errorbar(aes(ymin=sim_med-sim_sd,ymax=sim_med+sim_sd))+
  geom_point(data=all_fst_group,aes(x=emp_fst,y=sim_med))
```


## Check that, at the end of the simulation, there's still a reasonable spatial distribution of individuals

```{r}
## Read in and process the "individuals" .csv from tskit (but don't subset it to get just the sampled ones)
inds <- read_csv(file='../output/ts_8238198510834684392_t240000.csv',col_names=F,show_col_types=FALSE)
inds <- as.data.frame(t(inds))
inds <- separate_wider_delim(inds,cols=V1,delim=',',names=c('flags','x','y','zero','parents','array','dtype','pedID','pedp1','pedp2','age','subpop','sex','flags2')) %>%
  dplyr::select(x,y,pedID,pedp1,pedp2,age,sex)
inds <- mutate(inds,x=str_trim(str_sub(x,start=18)),pedID=str_trim(str_sub(pedID,-8,-1)),
               pedp1=str_trim(str_sub(pedp1,-8,-1)),pedp2=str_trim(str_sub(pedp2,-8,-1)),
               age=str_trim(str_sub(age,-1,-1)),sex=str_trim(str_sub(sex,-1,-1)))
inds <- mutate(inds,id=0:(nrow(inds)-1),x=as.numeric(x),y=as.numeric(y),age=as.numeric(age))

## Messy convert SLiM's x and y to lon and lat. (Messy is good enough, in this case)
load('../data/mmp_crop.RData') # use this to convert x and y into lat and lon
inds <- mutate(inds,lon = min_lon+x*(max_lon-min_lon)/dist_x, lat = min_lat+y*(max_lat-min_lat)/dist_y)

## Plot all individuals (in yellow) on top of the reference map
load('../data/TNC_sim.RData')
load('../data/SGS_sites.RData')
ggplot(barrier)+
  geom_sf(color='black',fill='black')+
  geom_sf(data=all_atolls,color='black',fill='black')+
  geom_sf(data=sgs_points_sf,color='blue')+
  coord_sf(xlim=c(min_lon,max_lon),ylim=c(min_lat,max_lat),expand=FALSE)+
  geom_sf_label(data=sgs_points_sf,aes(label=Site),color='blue',hjust=0, vjust=0,size=2)+
  theme(panel.background = element_rect(fill = "lightblue"),
        panel.grid.major = element_blank(),
        panel.border=element_blank(),
        plot.margin = unit(c(0, 0, 0, 0),"points"),
        panel.margin = unit(c(0, 0, 0, 0), "null"))+
  geom_point(data=inds,aes(x=lon,y=lat),color='yellow', size=0.01)
# save it as a really big png: e.g., 7070x6940, and zoom in, and you can see individuals distributed throughout the reef area.
```













## Scratch
Load a single empirical FST
```{r,eval=FALSE}
load(file=paste0(treefile,"_FST.RData"))
#load(file="../output/ts_4399080398350530317_t160000_FST.RData")
fst_sim <- as.data.frame(fsts) %>%
  rownames_to_column(var='site1')
rm(fsts)

fst_sim_long <- pivot_longer(fst_sim,cols=2:ncol(fst_sim),names_to=c('site2'))
fst_sim_long$sites <- NA
for(i in 1:nrow(fst_sim_long)){
  fst_sim_long$sites[i] <- paste(sort(as.matrix(fst_sim_long[i,1:2])),collapse='.')
}
fst_sim_long <- filter(fst_sim_long,!is.na(value)) %>%
  rename(sim_fst=value)
```

```{r}
all_fst_sim <- data.frame(sites=fst_sim_1$sites)
# add each fst dataset as a column
for(i in 1:length(fst_sims)){
  all_fst_sim <- full_join(all_fst_sim,fst_sims[[i]]) %>% select(-c(site1,site2))  
}

# calculate mean, median, sd of all the fst columns for each row (mutate)
just_data <- select(all_fst_sim,contains('sim_fst_'))
all_fst_sim$median <- apply(just_data,1,median)
all_fst_sim$sd <- apply(just_data,1,sd)
```

```{r}
fst_all <- full_join(fst_emp_long,fst_sim_long,by='sites') %>%
  select(sites,emp_fst,sim_fst) %>%
  filter(!is.na(emp_fst))

ggplot(data=fst_all,aes(x=emp_fst,y=sim_fst))+
  geom_point()+
  xlab('Empirical FST')+
  ylab('Simulated FST')
```