---
title: "Untitled"
author: "Your Name Here"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

setwd("C:/Users/eschlatter/Dropbox/DispersalSGS/analysis")
#treefile = "../output/ts_8238198510834684392_t160000"
```

Data prep:
```{r}
########## Load empirical FSTs

fst_emp <- read.csv('../data/FST_empirical.csv',header=T)
fst_emp_long <- pivot_longer(fst_emp,cols=2:ncol(fst_emp),names_to=c('site2'))
fst_emp_long$sites <- NA
for(i in 1:nrow(fst_emp_long)){
  fst_emp_long$sites[i] <- paste(sort(as.matrix(fst_emp_long[i,1:2])),collapse='.')
}
fst_emp_long <- filter(fst_emp_long,!is.na(value)) %>%
  rename(emp_fst=value)
fst_emp <- fst_emp_long
rm(fst_emp_long)


########## Load simulated FSTs

files <- list.files(path='../output/500k/FSTs') %>% strsplit(" ")
fst_sims <- list()
for (i in 1:length(files)) {
   load(paste0('../output/500k/FSTs/',files[[i]])) # load fst matrix

   fst_sim <- as.data.frame(fsts) %>%
     rownames_to_column(var='site1')
   fst_sim <- pivot_longer(fst_sim,cols=2:ncol(fst_sim),names_to=c('site2'))
   
   fst_sim$sites <- NA
   for(j in 1:nrow(fst_sim)){
     fst_sim$sites[j] <- paste(sort(as.matrix(fst_sim[j,1:2])),collapse='.')
   }
   fst_sim <- filter(fst_sim,!is.na(value)) %>%
     rename(sim_fst=value)
     #rename_with(~paste0('sim_fst_',i),value)
   
   assign(paste0('fst_sim_',i), fst_sim) # save as "fst_sim_i"
   fst_sims[[i]] <- fst_sim
}

########## Combine simulated and empirical

# combine simulation runs into a long table, with a column for the simulation number
all_fst_sim <- data.frame(site1=factor(), site2=factor(), sim_fst=numeric(), sites=factor(), sim=factor())
for(i in 1:length(fst_sims)){
  this_sim <- fst_sims[[i]]
  all_fst_sim <- rbind(all_fst_sim,mutate(this_sim, sim=as.factor(i)))
}

# add empirical data as its own column
all_fst_both <- full_join(all_fst_sim,fst_emp,by='sites') %>%
  select(sites,site1,site2=site2.x,sim,sim_fst,emp_fst) %>%
  filter(!is.na(emp_fst))

########## Group by site pair (across simulations)

all_fst_group <- group_by(all_fst_both,sites) %>%
  summarize(emp_fst=first(emp_fst),site1=first(site1),site2=first(site2),
            sim_med=median(sim_fst),sim_sd=sd(sim_fst),sim_min=min(sim_fst),sim_max=max(sim_fst))

# add extra columns for site, population, and population pair, to make exploratory plotting easier
site_list <- unique(c(fst_emp$X,fst_emp$site2))
site_list <- c(site_list[length(site_list)],site_list[1:(length(site_list)-1)]) # put them in order
for(i in site_list){
  all_fst_group <- mutate(all_fst_group,newsite=(i==site1 | i==site2)) %>%
    rename_with(~paste0(i),newsite)
}
for(i in c('B','T','G','L')){
  all_fst_group <- mutate(all_fst_group,newsite=grepl(i,sites)) %>%
    rename_with(~paste0(i),newsite)
}
all_fst_group <- mutate(all_fst_group,first_pop=substr(sites,1,1)) %>%
  mutate(second_pop=substr(sub('.*\\.','',sites),1,1)) %>%
  mutate(both_pops=paste0(first_pop,second_pop))

#save(all_fst_group,file='../output/FST_all_300k.RData')
```

Get pairwise distance between sites
```{r}
load('../output/FST_all_300k.RData')
sample_sites <- read_csv('../data/sgs_sites_SLiM',show_col_types=FALSE)

sitepair_dists <- dplyr::select(all_fst_group,sites,site1,site2)
sitepair_dists <- left_join(sitepair_dists,sample_sites,by=c('site1' ='Site'),keep=FALSE) %>%
  select(sites,site1,site2,site1.x=x,site1.y=y) %>%
  left_join(sample_sites,by=c('site2' ='Site'),keep=FALSE) %>%
  select(sites,site1,site2,site1.x,site1.y,site2.x=x,site2.y=y) %>%
  mutate(dist=sqrt((site1.x-site2.x)^2+(site1.y-site2.y)^2))

save(sitepair_dists, file='../data/sitepair_dists.RData')

```


```{r}
library(ggpattern)
load('../output/FST_all_300k.RData')

all_fst_group <- mutate(all_fst_group,both_pops_bt=case_match(both_pops,'BT'~'BB','GT'~'BG','LT'~'BL','TT'~'BB',.default=both_pops))
all_fst_group$both_pops_bt <- factor(all_fst_group$both_pops_bt,levels=c('BB','GG','LL','BG','BL','GL'))

#'#fc8d59' -orange (G)
#'#ffffbf' -yellow (L)
#'#91bfdb' -blue (B)

ggplot(data=all_fst_group,aes(x=sim_med,fill=both_pops_bt,pattern_fill=both_pops_bt,pattern_density=both_pops_bt))+
  geom_density_pattern(pattern_spacing=0.05,alpha=0.6,pattern_alpha=0.6)+
  scale_pattern_density_manual(values=c(0,0,0,0.5,0.5,0.5),name='Populations')+
  scale_fill_manual(values=c('#91bfdb','#fc8d59','#ffffbf','#91bfdb','#91bfdb','#fc8d59'),name='Populations')+
  scale_pattern_fill_manual(values=c('#91bfdb','#fc8d59','#ffffbf','#fc8d59','#ffffbf','#ffffbf'),name='Populations')+
  theme_bw()+
    theme(legend.position='bottom')+
  xlab('Simulated FST')+
  guides(pattern_density = guide_legend(nrow = 3))

ggplot(data=all_fst_group,aes(x=sim_med,fill=both_pops_bt,pattern_fill=both_pops_bt,pattern_density=both_pops_bt))+
  geom_density_pattern(pattern_spacing=0.03,alpha=0.6,pattern_alpha=0.6)+
  scale_pattern_density_manual(values=c(0,0,0,0.5,0.5,0.5),name='Populations')+
  scale_fill_manual(values=c('#91bfdb','#fc8d59','#ffffbf','#91bfdb','#91bfdb','#fc8d59'),name='Populations')+
  scale_pattern_fill_manual(values=c('#91bfdb','#fc8d59','#ffffbf','#fc8d59','#ffffbf','#ffffbf'),name='Populations')+
  theme_bw()+
  theme(legend.position='none', panel.grid.major = element_blank(),panel.grid.minor=element_blank())+
  labs(title='Simulated',x=NULL,y=NULL)

ggplot(data=all_fst_group,aes(x=emp_fst,fill=both_pops_bt,pattern_fill=both_pops_bt,pattern_density=both_pops_bt))+
  geom_density_pattern(pattern_spacing=0.03,alpha=0.6,pattern_alpha=0.6)+
  scale_pattern_density_manual(values=c(0,0,0,0.5,0.5,0.5),name='Populations')+
  scale_fill_manual(values=c('#91bfdb','#fc8d59','#ffffbf','#91bfdb','#91bfdb','#fc8d59'),name='Populations')+
  scale_pattern_fill_manual(values=c('#91bfdb','#fc8d59','#ffffbf','#fc8d59','#ffffbf','#ffffbf'),name='Populations')+
  theme_bw()+
  theme(legend.position='none', panel.grid.major = element_blank(),panel.grid.minor=element_blank())+
  labs(title='Empirical',x=NULL,y=NULL)
```

```{r}
library(ggpattern)
library(gridExtra)
load('../output/FST_all_300k.RData')

all_fst_group <- mutate(all_fst_group,both_pops_bt=case_match(both_pops,'BT'~'BB','GT'~'BG','LT'~'BL','TT'~'BB',.default=both_pops))
all_fst_group$both_pops_bt <- factor(all_fst_group$both_pops_bt,levels=c('BB','GG','LL','BG','BL','GL'))

bb <- ggplot(data=filter(all_fst_group,both_pops=="BB"),aes(x=sim_med))+
  geom_density_pattern(pattern_spacing=0.02,pattern_density=0,fill='#91bfdb',pattern_fill='#91bfdb')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.25)+
  xlab(NULL)+
  ylab(NULL)

gg <- ggplot(data=filter(all_fst_group,both_pops=="GG"),aes(x=sim_med))+
  geom_density_pattern(pattern_spacing=0.02,pattern_density=0,fill='#fc8d59',pattern_fill='#fc8d59')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.25)+
  xlab(NULL)+
  ylab(NULL)

ll <- ggplot(data=filter(all_fst_group,both_pops=="LL"),aes(x=sim_med))+
  geom_density_pattern(pattern_spacing=0.02,pattern_density=0,fill='#ffffbf',pattern_fill='#ffffbf')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.25)+
  xlab(NULL)+
  ylab(NULL)

bg <- ggplot(data=filter(all_fst_group,both_pops=="BG"),aes(x=sim_med))+
  geom_density_pattern(pattern_spacing=0.2,pattern_density=0.5,fill='#91bfdb',pattern_fill='#fc8d59')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.25)+
  xlab(NULL)+
  ylab(NULL)

bl <- ggplot(data=filter(all_fst_group,both_pops=="BL"),aes(x=sim_med))+
  geom_density_pattern(pattern_spacing=0.2,pattern_density=0.5,fill='#91bfdb',pattern_fill='#ffffbf')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.25)+
  xlab(NULL)+
  ylab(NULL)

gl <- ggplot(data=filter(all_fst_group,both_pops=="GL"),aes(x=sim_med))+
  geom_density_pattern(pattern_spacing=0.2,pattern_density=0.5,fill='#fc8d59',pattern_fill='#ffffbf')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text.y=element_blank(),axis.ticks.y=element_blank())+
  xlim(0,0.25)+
  xlab(NULL)+
  ylab(NULL)

g <- grid.arrange(bb,gg,ll,bg,bl,gl,ncol=1,
             left='Population pair',bottom='Simulated FST')
```
```{r}
library(ggpattern)
library(gridExtra)
load('../output/FST_all_300k.RData')

all_fst_group <- mutate(all_fst_group,both_pops_bt=case_match(both_pops,'BT'~'BB','GT'~'BG','LT'~'BL','TT'~'BB',.default=both_pops))
all_fst_group$both_pops_bt <- factor(all_fst_group$both_pops_bt,levels=c('BB','GG','LL','BG','BL','GL'))

bb <- ggplot(data=filter(all_fst_group,both_pops=="BB"),aes(x=emp_fst))+
  geom_density_pattern(pattern_spacing=0.02,pattern_density=0,fill='#91bfdb',pattern_fill='#91bfdb')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.03)+
  xlab(NULL)+
  ylab(NULL)

gg <- ggplot(data=filter(all_fst_group,both_pops=="GG"),aes(x=emp_fst))+
  geom_density_pattern(pattern_spacing=0.02,pattern_density=0,fill='#fc8d59',pattern_fill='#fc8d59')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.03)+
  xlab(NULL)+
  ylab(NULL)

ll <- ggplot(data=filter(all_fst_group,both_pops=="LL"),aes(x=emp_fst))+
  geom_density_pattern(pattern_spacing=0.02,pattern_density=0,fill='#ffffbf',pattern_fill='#ffffbf')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.03)+
  xlab(NULL)+
  ylab(NULL)

bg <- ggplot(data=filter(all_fst_group,both_pops=="BG"),aes(x=emp_fst))+
  geom_density_pattern(pattern_spacing=0.2,pattern_density=0.5,fill='#91bfdb',pattern_fill='#fc8d59')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.03)+
  xlab(NULL)+
  ylab(NULL)

bl <- ggplot(data=filter(all_fst_group,both_pops=="BL"),aes(x=emp_fst))+
  geom_density_pattern(pattern_spacing=0.2,pattern_density=0.5,fill='#91bfdb',pattern_fill='#ffffbf')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_blank(),axis.ticks=element_blank())+
  xlim(0,0.03)+
  xlab(NULL)+
  ylab(NULL)

gl <- ggplot(data=filter(all_fst_group,both_pops=="GL"),aes(x=emp_fst))+
  geom_density_pattern(pattern_spacing=0.2,pattern_density=0.5,fill='#fc8d59',pattern_fill='#ffffbf')+
  theme_bw()+
  theme(panel.border = element_blank(),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text.y=element_blank(),axis.ticks.y=element_blank())+
  xlim(0,0.03)+
  xlab(NULL)+
  ylab(NULL)

g <- grid.arrange(bb,gg,ll,bg,bl,gl,ncol=1,
             left='Population pair',bottom='Empirical FST')
```


Again, but let's do subsets
```{r}
load('../output/FST_all_300k.RData')
sample_sites <- read_csv('../data/sgs_sites_SLiM',show_col_types=FALSE)

fsts <- dplyr::select(all_fst_group,sites,site1,site2,emp_fst,sim_med) %>%
#  filter(!(site1 %in%c('B1','B2','B3'))&!(site2 %in%c('B1','B2','B3'))) %>%
  left_join(sample_sites,by=c('site1'='Site')) %>%
  dplyr::select(sites,site1,site2,emp_fst,sim_med,region1=Region) %>%
  left_join(sample_sites,by=c('site2'='Site')) %>%
  dplyr::select(sites,site1,site2,emp_fst,sim_med,region1,region2=Region) %>%
  mutate(region1=case_match(region1,
                            'barrier (northern)'~'N','barrier (central)'~'C',
                            'Turneffe Atoll'~'T',"Glover's Atoll" ~ 'G','Lighthouse Atoll'~'L',
                            .default='G')) %>%
  mutate(region2=case_match(region2,
                            'barrier (northern)'~'N','barrier (central)'~'C',
                            'Turneffe Atoll'~'T',"Glover's Atoll" ~ 'G','Lighthouse Atoll'~'L',
                            .default='G'))

fsts_subset <- filter(fsts,(region1 %in% c('N','C','T'))&(region2 %in% c('N','C','T')))

fsts_subset$regions <- NA
for(j in 1:nrow(fsts_subset)){
  fsts_subset$regions[j] <- paste(sort(as.matrix(fsts_subset[j,c('region1','region2')])),collapse='')
  }

fsts_subset$regions <- factor(fsts_subset$regions,levels=c('NN','CC','TT','CN','NT','CT'))

#'#2b8cbe' -northern (N)
#'#a6bddb' -central (C)
#'#ece7f2' -turneffe (T)

ggplot(data=fsts_subset,aes(x=sim_med,fill=regions,pattern_fill=regions,pattern_density=regions))+
  geom_density_pattern(pattern_spacing=0.05,alpha=0.5,pattern_alpha=0.5)+
  scale_pattern_density_manual(values=c(0,0,0,0.5,0.5,0.5),name='Regions')+
  scale_fill_manual(values=c('#2b8cbe','#a6bddb','#ece7f2','#2b8cbe','#2b8cbe','#a6bddb'),name='Regions')+
  scale_pattern_fill_manual(values=c('#2b8cbe','#a6bddb','#ece7f2','#a6bddb','#ece7f2','#ece7f2'),name='Regions')+
  theme_bw()+
  xlab('Simulated FST')+
    theme_bw()+
    theme(legend.position='bottom')+
  xlab('Simulated FST')+
  guides(pattern_density = guide_legend(nrow = 3))

ggplot(data=fsts_subset,aes(x=sim_med,fill=regions,pattern_fill=regions,pattern_density=regions))+
  geom_density_pattern(pattern_spacing=0.02,alpha=0.5,pattern_alpha=0.5)+
  scale_pattern_density_manual(values=c(0,0,0,0.5,0.5,0.5),name='Regions')+
  scale_fill_manual(values=c('#2b8cbe','#a6bddb','#ece7f2','#2b8cbe','#2b8cbe','#a6bddb'),name='Regions')+
  scale_pattern_fill_manual(values=c('#2b8cbe','#a6bddb','#ece7f2','#a6bddb','#ece7f2','#ece7f2'),name='Regions')+
  theme_bw()+
  theme(legend.position='none', panel.grid.major = element_blank(),panel.grid.minor=element_blank())+
  labs(title='Simulated',x=NULL,y=NULL)

ggplot(data=fsts_subset,aes(x=emp_fst,fill=regions,pattern_fill=regions,pattern_density=regions))+
  geom_density_pattern(bw=0.0005,pattern_spacing=0.02,alpha=0.6,pattern_alpha=0.6)+ # try bw=0.0005 to remove lumpiness
  scale_pattern_density_manual(values=c(0,0,0,0.5,0.5,0.5),name='Regions')+
  scale_fill_manual(values=c('#2b8cbe','#a6bddb','#ece7f2','#2b8cbe','#2b8cbe','#a6bddb'),name='Regions')+
  scale_pattern_fill_manual(values=c('#2b8cbe','#a6bddb','#ece7f2','#a6bddb','#ece7f2','#ece7f2'),name='Regions')+
  theme_bw()+
  ylim(0,1000)+
  theme(legend.position='none', panel.grid.major = element_blank(),panel.grid.minor=element_blank())+
  labs(title='Empirical',x=NULL,y=NULL)

```



Check IBD
```{r}
load('../output/FST_all_300k.RData')
load('../data/sitepair_dists.RData')

all_fst_group <- left_join(all_fst_group,sitepair_dists,keep=FALSE) %>%
  filter(both_pops=='BB')

all_fst_group <- mutate(all_fst_group,
                        emp_scale=emp_fst/(1-emp_fst),
                        sim_scale=sim_med/(1-sim_med))

ggplot(all_fst_group,aes(x=dist,y=sim_scale))+
  geom_point(size=1.1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(title='Simulated',x=NULL,y=NULL)
#  geom_smooth(method='lm',se=FALSE)+


ggplot(all_fst_group,aes(x=dist,y=sim_med))+
  geom_point(size=1.1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(title='Simulated',x=NULL,y=NULL)
#  geom_smooth(method='lm',se=FALSE)+


ggplot(all_fst_group,aes(x=dist,y=emp_scale))+
  geom_point(size=1.1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(title='Simulated',x=NULL,y=NULL)
#  geom_smooth(method='lm',se=FALSE)+

ggplot(all_fst_group,aes(x=dist,y=emp_fst))+
  geom_point(size=1.1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(title='Empirical',x=NULL,y=NULL)
```

Check IBD
```{r}
load('../output/FST_all_300k.RData')
load('../data/sitepair_dists.RData')

all_fst_group <- left_join(all_fst_group,sitepair_dists,keep=FALSE)

ggplot(all_fst_group,aes(x=dist,y=sim_med,color=(B1|B2|B3)))+
  facet_wrap(~both_pops,nrow=2)+
  geom_point(size=1.1)+
  theme_bw()+
  xlab('Distance (km)')+
  ylab('Simulated FST (median of 10 sims)')+
#  geom_smooth(method='lm',se=FALSE)+
  ggtitle('Simulated IBD by population pair')

ggplot(all_fst_group,aes(x=dist,y=emp_fst,color=(B1|B2|B3)))+
  facet_wrap(~both_pops,nrow=2)+
  geom_point(size=1.1)+
  theme_bw()+
  xlab('Distance (km)')+
  ylab('Empirical FST')+
#  geom_smooth(method='lm',se=FALSE)+
  ggtitle('Empirical IBD by population pair')
```

```{r}
load('../output/FST_all_300k.RData')

ggplot(all_fst_group,aes(x=emp_fst,y=sim_med))+
  geom_point()+
  theme_bw()+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')+
  ggtitle('4 chromosomes, 10k generations')
```



```{r}
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_point()+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')+
  geom_smooth(method='lm',se=FALSE)+
  ggtitle('10k generations, 4 chromosomes')
```

```{r}
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_point()+
  theme_bw()+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')+
  geom_smooth(method='lm',se=FALSE)+
  ggtitle('300k generations')
```


```{r}
library(cowplot)

load('../output/FST_all_20k.RData')
all_fst_group_20k <- all_fst_group
rm(all_fst_group)
load('../output/FST_all_120k.RData')
all_fst_group_120k <- all_fst_group
rm(all_fst_group)
load('../output/FST_all_180k.RData')
all_fst_group_180k <- all_fst_group
rm(all_fst_group)
load('../output/FST_all_300k.RData')
all_fst_group_300k <- all_fst_group
rm(all_fst_group)

g1 <- ggplot(all_fst_group_20k,aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_point()+
  theme_bw()+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')+
  geom_smooth(method='lm',se=FALSE)+
  ggtitle('20k generations')+
  theme(legend.position = "none")+
  ylim(-0.01,0.25)

g2 <- ggplot(all_fst_group_120k,aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_point()+
  theme_bw()+
  xlab('Empirical FST')+
  ylab(NULL)+
  geom_smooth(method='lm',se=FALSE)+
  ggtitle('120k generations')+
  theme(legend.position = "none")+
  ylim(-0.01,0.25)

g3 <- ggplot(all_fst_group_180k,aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_point()+
  theme_bw()+
  xlab('Empirical FST')+
  ylab(NULL)+
  geom_smooth(method='lm',se=FALSE)+
  ggtitle('180k generations')+
  theme(legend.position = "none")+
  ylim(-0.01,0.25)

g4 <- ggplot(all_fst_group_300k,aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_point()+
  theme_bw()+
  xlab('Empirical FST')+
  ylab(NULL)+
  geom_smooth(method='lm',se=FALSE)+
  ggtitle('300k generations')+
  theme(legend.position = "none")+
  ylim(-0.01,0.25)

grid.arrange(g1,g2,g3,g4,nrow=1)

```


```{r}
ggplot(filter(all_fst_group,both_pops %in% c('TT','GG','LL')),aes(x=emp_fst,y=sim_med,color=both_pops))+
  geom_pointrange(aes(ymin=sim_min,ymax=sim_max,color=both_pops),position=position_jitter(width=0.0001))+
  xlab('Empirical FST')+
  ylab('Simulated FST (median of 10 sims)')+
  geom_smooth(method='lm',se=FALSE)
```


Looks like there's something going on with the Barrier-Lighthouse group. Some simulated values are much larger than others, for similar empirical values.
First check: is it just among-sim variation?
```{r}
BL <- filter(all_fst_group,both_pops=="BL")# %>% select(sites,emp_fst,sim_med,sim_sd,sim_min,sim_max)
ggplot(BL,aes(x=emp_fst,y=sim_med))+
  geom_pointrange(aes(ymin=sim_med-sim_sd,ymax=sim_med+sim_sd),
                  position=position_jitter(width=0.0008))+
  geom_point(data=all_fst_group,aes(x=emp_fst,y=sim_med),color='gray')
```
Nope. There are definitely some points that are distinct from the rest. Figure out which ones they are:
It's B1, B2, and B3 with Lighthouse.
```{r}
filter(BL,sim_med>0.175)

## highlight the points from one site
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med,fill=both_pops,color=both_pops),shape=21)+
  theme_bw()+
  geom_point(size=1.1)+
  geom_smooth(method='lm',se=FALSE)+
  geom_point(data=filter(all_fst_group,
                         (B1 | B2 | B3)),
             aes(x=emp_fst,y=sim_med,fill=both_pops),color='black',shape=21)+
  ggtitle('B1,B2,B3')
```

```{r}
## highlight a single point
ggplot(all_fst_group,aes(x=emp_fst,y=sim_med))+
  geom_point(size=1)+
  geom_point(data=filter(all_fst_group,B10==TRUE & T6==TRUE),aes(x=emp_fst,y=sim_med),color='red')
```


## Check that, at the end of the simulation, there's still a reasonable spatial distribution of individuals

```{r}
sample_sites <- read_csv('../data/sgs_sites_SLiM',show_col_types=FALSE)

## Read in and process the "individuals" .csv from tskit (but don't subset it to get just the sampled ones)
inds <- read_csv(file='../output/sample_inds/ts_8238198510834684392_t300000.csv',col_names=F,show_col_types=FALSE)
inds <- as.data.frame(t(inds))
inds <- separate_wider_delim(inds,cols=V1,delim=',',names=c('flags','x','y','zero','parents','array','dtype','pedID','pedp1','pedp2','age','subpop','sex','flags2')) %>%
  dplyr::select(x,y,pedID,pedp1,pedp2,age,sex)
inds <- mutate(inds,x=str_trim(str_sub(x,start=18)),pedID=str_trim(str_sub(pedID,-8,-1)),
               pedp1=str_trim(str_sub(pedp1,-8,-1)),pedp2=str_trim(str_sub(pedp2,-8,-1)),
               age=str_trim(str_sub(age,-1,-1)),sex=str_trim(str_sub(sex,-1,-1)))
inds <- mutate(inds,id=0:(nrow(inds)-1),x=as.numeric(x),y=as.numeric(y),age=as.numeric(age))

## Messy convert SLiM's x and y to lon and lat. (Messy is good enough, in this case)
load('../data/mmp_crop.RData') # use this to convert x and y into lat and lon
inds <- mutate(inds,lon = min_lon+x*(max_lon-min_lon)/dist_x, lat = min_lat+y*(max_lat-min_lat)/dist_y)

site_coords <- sample_sites[,c('Site','x','y')]
site_coords <- mutate(site_coords,lon = min_lon+x*(max_lon-min_lon)/dist_x, lat = min_lat+y*(max_lat-min_lat)/dist_y)

## Plot all individuals (in yellow) on top of the reference map
load('../data/TNC_sim.RData')
load('../data/SGS_sites.RData')
ggplot(barrier)+
  geom_sf(color='black',fill='black')+
  geom_sf(data=all_atolls,color='black',fill='black')+
  geom_sf(data=sgs_points_sf,color='blue')+
  coord_sf(xlim=c(site_coords[site_coords$Site=='B3',]$lon-0.12,site_coords[site_coords$Site=='B3',]$lon+0.15),
           ylim=c(site_coords[site_coords$Site=='B3',]$lat-0.3,max_lat),expand=FALSE)+
  #coord_sf(xlim=c(min_lon,max_lon),ylim=c(min_lat,max_lat),expand=FALSE)+
  geom_sf_label(data=sgs_points_sf,aes(label=Site),color='blue',hjust=0, vjust=0,size=6)+
  theme(panel.background = element_rect(fill = "lightblue"),
        panel.grid.major = element_blank(),
        panel.border=element_blank(),
        plot.margin = unit(c(0, 0, 0, 0),"points"),
        panel.margin = unit(c(0, 0, 0, 0), "null"))+
  geom_point(data=inds,aes(x=lon,y=lat),color='yellow', size=0.005)

# save it as a really big png: e.g., 7070x6940, and zoom in, and you can see individuals distributed throughout the reef area.
```